\section{Monday for MAT3006}\index{Monday_lecture}

\subsection{Arzela-Ascoli Theorem}
\begin{theorem}
Let $\mathcal{F}$ be a closed set in $\mathcal{C}(K)$, where $K$ is closed and bounded in $\mathbb{R}^n$.
Then $\mathcal{F}$ is compact iff it is bounded and equi-continuous.
\end{theorem}
\begin{definition}[Equi-continuous]
For any $\varepsilon>0$, there exists $\delta>0$ such that whenever $d(x,y)<\delta, x,y\in X$
\[
d_\infty(f(x),f(y))<\varepsilon,\forall f\in\mathcal{F}
\]
\end{definition}
\begin{example}
Let $X$ be a bounded and convex set in $\mathbb{R}^n$.
Show that a family of equi-continuous functions, say $\mathcal{F}$,
is bounded in $C(X)$,
if there exists a point $x_0\in X$ and a constant $M>0$ such that $|f(x_0)|\le M$ for all $f\in\mathcal{F}$.
\begin{proof}
By equi-continuity for $\varepsilon:=1$, there exists $\delta_0$ such that $|f(x)-f(y)|<1$ whenever $|x-y|<\delta_0$.

Consider the open ball $B_{R}(x_0)$ containing $E$.
Then $|x-x_0|\le R$ for all $x\in E$, we can find $x_0,x_1,\dots,x_n=x$ where $n\delta_0\le R\le (n+1)\delta_0$ such taht
\[
|x_{n+1} - x_0|\le \delta_0\implies
|f(x)-f(x_0)|\le \sum_{j=0}^{n-1}|f(x_{j+1}) - f(x_j)|\le n\le R/\delta_0
\]
Therefore,
\[
|f(x)|\le |f(x_0)|+n+1\le M+R/\delta_0,\forall x\in X,\forall f\in\mathcal{F}
\]
\end{proof}
\end{example}

\begin{example}
Let $\phi:[0,1]\times\mathbb{R}\to\mathbb{R}$ be continuous. Suppose that there is $M>1$ such that 
$|\phi(r,s)|\le M$ for $r\in[0,1]$ and $s\in\mathbb{R}$.
Let $\mathbb{Z} = C[0,1]$ and $\psi$ be the following:
\[
\psi(f(x)) = c + \int_0^x\phi(t,f(t))\diff t,\forall x\in[0,1]
\]
where $c$ is a constant.
Then define
\[
E = \left\{f\in\mathbb{Z}\middle|
|f(x)-c|\le M,\quad
|f(x)-f(y)|\le M|x-y|,\ \forall x,y\in[0,1]
\right\}
\]
Show that $\psi(E)\subseteq E$ and $E$ is compact.
\end{example}
\begin{proof}
Let $f\in E$, then
\[
|\psi(f(x)) - c|=|\int_0^x\phi(t,f(t))\diff t|\le 1\cdot M = M
\]
Moreover,
\[
|\psi(f(x)) - \psi(f(y))|=|\int_y^x\phi(t,f(t))\diff t|\le M|x-y|=M|x-y|
\]
Therefore, $\psi(E)\subseteq E$.

We need to show that $E$ is bounded and equi-continuous:

Since $|f(x)-c|\le M,\forall f\in E$, we imply $|f(x)|\le M+c$.

Given $\varepsilon>0$, take $\delta = \frac{\varepsilon}{M}$, then 
\[
|x-y|<\delta\implies|f(x) - f(y)|\le M|x-y|=\varepsilon,\forall f\in E
\]

The closedness of $E$:
for any $\{f_n\}\in E$ and let $f_n\to f$, and show that $f\in E$, i.e., $E'\subseteq E$, i.e., $E$ is closed.

Applying Arzela-Ascoli Theorem, $E$ is compact.
\end{proof}

\subsection{Baire Category Theorem}
Let $\{E_j\}_{j=1}^\infty$ be a sequence of nowhere dense subsets of $(X,d)$, where $(X,d)$ is complete.
Then
\begin{itemize}
\item
$\cup_{j=1}^\infty \overline{E}_j$ has empty interior.
\end{itemize}

\begin{definition}[Nowhere Dense]
A set $E$ is said to be \emph{nowhere dense} if $X\setminus E$ is dense in $X$.
\end{definition}
Cator Set: nowhere dense.

Boundary point is nowhere dense.

\begin{proposition}
The intersection of dense sequence of open sets is dense.
\end{proposition}
\begin{proof}
Let $\{G_j\}$ be open and $\overline{G}_j = X$.
Let $F_j = X\setminus G_j$.
Then $F_j$ is nowhere dense, which follows that
\[
\bigcap_j G_j = \left(\bigcup_{j}F_j\right)^c,
\]
which is comeagre, and therefore dense.
\end{proof}
\begin{proposition}
The set $\mathbb{Q}$ of rationals is an $F_\delta$ but not $G_\delta$, where $F_\delta = \cup_{n=1}^NF_n$, where $F_n$ is closed, and $G_\delta = \cap_{n=1}^NG_n$, where $G_n$ is open.
\end{proposition}
\begin{proof}
If $\mathbb{Q}$ were a $G_\delta$, then the complement of $\mathbb{Q}$ is a countable union of closed sests 
\[
\mathbb{Q}^c=\cup_{n=1}^NG_n^c
\]
Since $\mathbb{Q}^c$ has no interiors, $\text{int}(\mathbb{Q}^c)$.
However, $\mathbb{R} = \mathbb{Q}\cup(\cup_{n=1}^NG_n^c)$, a countable union of nowhere dense sets, which contradicts to the BCT.
\end{proof}















