\section{Wednesday for MAT4002}\index{Monday_lecture}
Triangulation of topological space $X$
\[
V=\{1,2,3,4,\dots,9\},\quad
\Sigma=\left\{
\begin{aligned}
\text{$9$ subsets with $1$ element}\\
\text{$27$ subsets with $2$ elements}\\
\text{$18$ subsets with $3$ elements}
\end{aligned}
\right.
\]

Figures

Question: can we go from a figure into $(V,\Sigma)?$
e.g.,
the figure is not a triangulation.

Now we can do the following:
\[
\text{study of $X$}\mapsto\{\text{study of $(V,\Sigma)$ with $|(V,\Sigma)|\cong X$}\}
\]
e.g., can we distinguish $X=S^1\times S^1$ and $Y=S^2$, .e.g., can we distinguish $|(V_1,\Sigma_1)|$ and $|(V_2,\Sigma_2)|$?

\begin{theorem}
Suppose that $|(V_1,\Sigma_1)|\cong|(V_2,\Sigma_2)|$, then 
\begin{align*}
&\sum_{i=1}^\infty(-1)^i\left(\text{number of subsets in $\Sigma_1$ with $(i+1)$-element}\right)\\
&=
\sum_{i=1}^\infty(-1)^i\left(\text{number of subsets in $\Sigma_2$ with $(i+1)$-element}\right)
\end{align*}
This goes back to Euler: $V-E+F = 2$.
\end{theorem}
Since $\mathcal{X}(S^2) = 5-9+6=2$ and $\mathcal{X}(S^1\times S^1) = 9-27+18=0$, we conclude that
\[
S^2\not\cong S^1\times S^1,
\]

\subsection{Simplicial Sub-complex}

\begin{definition}
A subcomplex of a simplicial complex $K=(V,\Sigma)$ is a simplicial complex $K'=(V',\Sigma')$ with 
\[
V'\subseteq V,\quad
\Sigma'\subseteq\Sigma
\]
\end{definition}
\begin{proposition}
Suppose $K'$ is subcomplex of $K$, then $|K'|$ is closed in $|K|$.
\end{proposition}
\begin{proof}
Suppose that $D$ is the disjoint union of all the simplicial complex forming $|K|$. (number of component in $D$ is $|\Sigma|$)
and
$p:D\to|K|$ be the projection mapping.

Note that $p^{-1}(|K'|)$ are precisely the (disjoint union of) $\Delta_{\sigma'}$, where $\sigma'\in\Sigma'$, which is closed in $D$.
By definition of quotient topologu, $|K'|$ is also closed. 
\end{proof}

\begin{definition}[sub-complex]
Let $K=(V,\Sigma)$ be simplicial complex and $V'\subseteq V$.
Then the subcomplex spanned by $V'$ has
\begin{itemize}
\item
vertex $V'$
\item
simplices $\Sigma'$ given by
\[
\{\sigma\in\Sigma\mid\sigma\subseteq V'\}
\]
\end{itemize}
\end{definition}

\begin{definition}[Link and Star]
Let $(V,\Sigma)=K$ be simplicial complex
\begin{itemize}
\item
The \emph{link} of $\bm v\in V$, denoted as $\text{lk}(\bm v)$ is the sub-complex with
\begin{itemize}
\item
verex set
\[
\{\bm w\in V\setminus\{\bm v\}\mid \{\bm v,\bm w\}\in\Sigma\}
\]
\item
simplicies
\[
\{\sigma\in\Sigma\mid\bm v\notin\sigma\text{ and }\sigma\cup\{\bm v\}\in\Sigma\}
\]
\end{itemize}
\item
The star of $\bm v$~($\text{st}(\bm v)$) is
\[
\bigcup\{\text{inside}(\sigma)\mid \sigma\in\Sigma,\bm v\in\sigma\}
\]
\end{itemize}
\end{definition}

\begin{proposition}
$\text{st}(\bm v)$ is open and $\bm v\in\text{st}(\bm v)$.
\end{proposition}
\begin{proof}
Omitted.
\end{proof}
In fact, $|K|\setminus\text{st}(\bm v)$ is the simplicial subcomplex spanned by $V$.

\subsection{Some properties of simplicial complex}

\begin{proposition}
Suppose that $K=(V,\Sigma)$, where $V$ is finite.
Then $|K|$ is compact.
\end{proposition}
\begin{proof}
The mapping $p:D\to|K|$ is a projection mapping; and 
$D$ (as a finite disjoint union of $\Delta_\sigma$'s) is compact.

Therefore, $p(D)=|K|$ is compact.
\end{proof}

\begin{proposition}
For each $K=(V,\Sigma)$, where $V$ is finite, there is a continuous injection
\[
f:|K|\to\mathbb{R}^n\ \text{ for some $n$}
\]
\end{proposition}
\begin{proof}
Let $K'=(V,\Sigma')$, where $\Sigma'=\text{power set of $V$}$. Then
\[
|K'| = \Delta^{|V|-1}\subseteq \mathbb{R}^{|V|}
\]
Consider
\[
i:|K|\to|K'|
\]
which comes from the following
\[
(V,\Sigma)\to D=\sqcup_{\sigma\in \Sigma}\Delta_{\sigma}
\]
\[
(V,\Sigma')\to D'=\sqcup_{\sigma'\in \Sigma'}\Delta_{\sigma'}
\]
$\tilde{i}:D\to D'\to_{p'}|K|$, which descends to the continuous injective
\[
i:D/n\to|K'|
\]
\[
|K|\to^i|K'|=\Delta^{|V|-1}\to\mathbb{R}^{|V|}
\]
i.e.., $|K|\to\mathbb{R}^{|V|}$.
\end{proof}














