\section{Wednesday}

\begin{proposition}
$G$ is hamiltionian implies that for any nonempty $S\subseteq V$, $G-S$ has at most $|S|$ components.
\end{proposition}
\begin{theorem}
$G$ is Hamiltionian if for any vetices $v,w$ such that $(v,w)\notin E$,
\[
\mbox{deg}(v)+\mbox{deg}(w)\ge n
\]
\end{theorem}


\paragraph{Knapsacle Problem}
\[
\begin{array}{ll}
\max&\sum_{i=1}^nv_iu_i\\
\mbox{such that}&\sum_{i=1}^nw_iu_i\le W\\
&u_i\in\{0,1\},\ i=1,\dots,n
\end{array}
\]

This is the binary integer programming problem, which is NP-complete. We cannot find exact solution to this problem. There are $2^n$ possible solutions, which requires exponential time.

We can find some ``pseudo''-polynomial algorithm. Assumption:
\begin{enumerate}
\item
$W$ is an integer. 

\end{enumerate}

State: remaining capacity, define as $X_k$.

Stage: iterm $1,\dots,n$.
\[
J_N(x_N)=\left\{
\begin{aligned}
v_N &\quad\text{if }N\le X_N\\
0,&\quad\text{if }w_N>X_N
\end{aligned}
\right.
\]
\[
J_N(x_{N-1})
=
\left\{
\begin{aligned}
0+J_N(X_N),&\quad\text{if }w_{N-1}>X_{N-1}, \text{Here }X_N=X_{N-1}\\
\max\{v_{N-1}+J_N(x_N), 0+J_N(x_N)\}&,\text{ if }w_{N-1}\le X_{N-1}
\end{aligned}
\right.
\]

Here why the DP has ``pesduo''-polynomial time, and the $w_i$ should be integer?

Let's list states $\{x_1,\dots,x_N\}$. For each stage $k$, there would be $W+1$.

Each iteration at most $2$ computation, and therefore we face $2(W+1)\cdot N$

\subsection{Label Correcting Methods}
Shortest Path. Back up some information.

The label refers to the intermediate information.

$A^*$-algorithm; Bellman-Ford Algorithm

Benchmark: MILP, CPLEX, CVX.

\subsection{DP problems with perfect state information}

\paragraph{Linear Quadratic System}
Let $x_k\in\mathbb{R}^1$ be the state; $u_k\in\mathbb{R}^n$ be the control; $\omega_k\in\mathbb{R}^n$ be the disturbance.

System Dynamics.
\[
x_{k+1}=A_kx_k+B_ku_k+\omega_k
\]
Stage cost:
\[
x_k'Q_kx_k+u_k'R_ku_k
\]
Here $Q_k$ is required to be positive-semi-definite symmetric matrix; $R_k$ to be positive-definite symmetric matrix.
\[
J_{n-1}(x_{n-1})=\min_{u_{n-1}}\mathbb{E}_{\omega}
\left\{
x_{n-1}'Q_{n-1}x_{n-1}+u_{n-1}'R_{n-1}u_{n-1}
+
J_n(x_n)
\right\}
\]
where
\[
J_n(x_n)=(A_{n-1}x_{n-1}+Bu_{n-1}+\omega)'Q_n
(A_{n-1}x_{n-1}+Bu_{n-1}+\omega)
\]













